Cycle Route Uptake and Scenario Estimation (CRUSE): an approach for developing strategic cycle network planning tools

### 2024-03-26 paper.qmd
add edits from Kildare

Thoughout the project, we undertook a series of workshops and interviews with stakeholders in Ireland to gather feedback on the approach and resulting tool.
The feedback was generally positive, with stakeholders commenting on the need for the tool, lack of data on cycling potential proving a barrier to decision-making and investment, and the potential for the tool.
At an end of project workshop in December 2023, stakeholders were asked to provide feedback on the tool and its potential use in practice.
Quotes and suggestions from these workshops, and a subsequent interview with a Senior Transport Planning using the evidence in their everyday work in Transport Infrastructure Ireland, and presented in this section (with quotes being attributed to named individuals where permission has been granted).

Early in the project, we presented the approach to practitioners in Kildare and Limmerick.
The feedback from the workshop in Kildare highlighted the need for the tool for practitioners at the county level, as highlighted in the following quote:

> "It's a missing piece of evidence that will help new projects get built" (Senior Engineer, Sustainable Transport & Traffic Management, Kildare County Council)

Practitioners in Kildare County Council that the evidence generated *was already valuable* the development of new cycle routes, especially when making the case during planning applications. They emphasized that cycle networks are often fragmented, and the tool allows them to demonstrate how individual local schemes can contribute to a larger, interconnected network. Council officers expressed a preference for visualizations that directly compare the fast and quiet routes, as they recognize the importance of both types of routes (users can switch between the on the 'Route types' page for each county although this functionality is not available in the landing page map due to other feedback requesting a simpler interface).

In comparison with another tool (the National Transport Authority's cycling propensity tool), the CRUSE tool was seen by practitioners in Kildare as more user-friendly and intuitive, with the ability to generate evidence at the route level, which is crucial for planning new cycling infrastructure.
Overall, the workshop participants concluded that the approach filled gap in evidence and will greatly assist in the development of new cycling projects.

@@

In a subsequent interview with a practitioner using the evidence generated by the CRUSE approach in their everyday work, we received the following feedback.
The outputs were a "great tool for getting started with project outline documents" and "ideal for generating a strategic rationale for estimating demand and building the business case in the absence of cycle count data" (Declan Keenan, Senior Transport Planner working on active travel, personal communication).
Declan also noted the alignment between local knowledge, cycle count data, and the outputs of the approach: "It's a good representation of what's going on" and the networks allign well in terms of network "shape" and the proportion of trips of trips on different parts of the network.
There were two suggestions from Declan which have not yet been implemented: the provision of data showing only "gold standard" POWSCAR data, and the need for training on the tool to ensure that all staff can use it effectively.
These, and other potential improvements, are discussed in the next section.

>>>>

Throughout the project, we undertook a series of workshops and interviews with stakeholders in Ireland to gather feedback on the approach and resulting tool.
The feedback was generally positive, with stakeholders commenting on the need for the tool, lack of data on cycling potential proving a barrier to decision-making and investment, and the potential for the tool.
At an end of project workshop in December 2023, stakeholders were asked to provide feedback on the tool and its potential use in practice.
Quotes and suggestions from these workshops, and a subsequent interview with a Senior Transport Planning using the evidence in their everyday work in Transport Infrastructure Ireland, and presented in this section (with quotes being attributed to named individuals where permission has been granted).

Early in the project, we presented the approach to practitioners in Kildare and Limerick.
The feedback from the workshop in Kildare highlighted the need for the tool for practitioners at the county level, as highlighted in the following quote:

> "It's a missing piece of evidence that will help new projects get built" (Dónal Hodgins, Senior Engineer, Sustainable Transport & Traffic Management, Kildare County Council)

Practitioners in Kildare County Council found that the evidence generated *was already valuable* the development of new cycle routes, especially when making the case during planning applications. They emphasized that cycle networks are often fragmented, and the tool allows them to demonstrate how individual local schemes can contribute to a larger, interconnected network. Council officers expressed a preference for visualizations that directly compare the fast and quiet routes, as they recognize the importance of both types of routes (users can switch between the on the 'Route types' page for each county although this functionality is not available in the landing page map due to other feedback requesting a simpler interface).

In comparison with another cycling propensity tool, the CRUSE Tool was seen by practitioners in Kildare as more user-friendly and intuitive, with the ability to generate evidence at the route level, which is crucial for planning new cycling infrastructure.
Overall, the workshop participants concluded that the approach filled a gap in evidence and will greatly assist in the development of new cycling projects.

@@

In a subsequent interview with a practitioner using the evidence generated by the CRUSE approach in their everyday work, we received the following feedback.
The outputs were a "great tool for getting started with project outline documents" and "ideal for generating a strategic rationale for estimating demand and building the business case in the absence of cycle count data" (Declan Keenan, Senior Transport Planner working on active travel, personal communication).
Declan also noted the alignment between local knowledge, cycle count data, and the outputs of the approach: "It's a good representation of what's going on" and the networks align well in terms of network "shape" and the proportion of trips of trips on different parts of the network.
There were two suggestions from Declan which have not yet been implemented: the provision of data showing only "gold standard" POWSCAR data, and the need for training on the tool to ensure that all staff can use it effectively.
These, and other potential improvements, are discussed in the next section.
